Learning constraints in tabular data

### 2016-05-20 Paper/ECAI/tabular.tex
Reworked superblock assignment

\usepackage{times}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{xspace}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[numbers]{natbib}
\usepackage{notoccite}
\usepackage{fixltx2e}
\usepackage{framed}
\usepackage{amsmath}
\usepackage{tabularx}
\usepackage{subcaption}
\usepackage[inline]{enumitem}

@@

\newcommand{\CName}{Syntax\xspace}
\newcommand{\CSignature}{Signature\xspace}
\newcommand{\CFunction}{Definition\xspace}
\newcommand{\dependencies}{\ensuremath{\mathcal{D}}\xspace}
\newcommand{\groups}{\ensuremath{\mathcal{G}}\xspace}

@@

\newcommand{\numeric}{\format{numeric}}
\newcommand{\textual}{\format{textual}}
\newcommand{\integer}{\format{integer}}
\newcommand{\discrete}{\format{discrete}}
\newcommand{\plength}{\format{length}}
\newcommand{\ptype}{\format{type}}
\newcommand{\ptable}{\format{table}}
\newcommand{\por}{\format{orientation}}
\newcommand{\prows}{\format{rows}}
\newcommand{\pcols}{\format{columns}}
\newcommand{\nat}{\mathcal{N}}

\newcommand{\sg}{B}

@@

The question that we answer in this paper is: is it possible to discover or reconstruct structural constraints (relations, functions) in flat tabular spreadsheet data?
To answer this question we contribute a general-purpose method and system, named \sname (pronounced as the word ``tackle''), for discovering row-wise and column-wise constraints.
%Furthermore, the algorithm accepts declarative specifications of the constraints and predicates of interest (the "bias"). \tias{revise declarative claim?}
It operates directly on (headerless) tables of a spreadsheet in an unsupervised setting, as it reasons on that raw tabular data directly with no example constraint instantiations given. %, and discovers column- and row-wise constraints.
We demonstrate the utility of our approach in an experimental evaluation.
Moreover, we sketch additional application scenario's such as autocompletion and error detection.
\\\\
This paper is organized as follows.
Section \ref{sec:formalization} introduces concepts relevant to our approach. Section \ref{sec:problem_statement} presents the problem statement and the approach. Section \ref{sec:evaluation} presents the evaluation of the approach. Section \ref{sec:applications} shows how our system can be used for applications. Section \ref{sec:related_work} presents an overview of the related work. Section \ref{sec:conclusions} provides conclusions.

@@

A row or a column is \textbf{type-consistent} if all cells in that row or column are of the same base type, that is, numeric or textual. We will use notation $T[a,{:}]$ to refer to the $a$-th row of table $T$, and similarly $T[{:},a]$ for the $a$-th column. For example in Figure~\ref{fig:main_example}, $T_1[1,:] = [1,2,3,4]$ and $T_5[:,1] = ['$Diana Coolen$', 5]$. The latter is not type-consistent while the former is.
%We will write \textbf{vector} to refer to a single row or column when its orientation does not matter.

@@

\begin{definition}
\textbf{Block containement $\sqsubseteq$.} 
A block $B'$ is contained in a block $B$, $B' \sqsubseteq B$, iff both are valid blocks (single orientation, contiguous, type consistent) and each of the vectors in $B'$ is also in $B$. For row-oriented blocks: $B' \sqsubseteq B \Leftrightarrow B=T[\rangeto{a}{b},:] \wedge B'=T[\rangeto{a'}{b'},:] \wedge a \leq a' \wedge b' \leq b$ and similarly for column-oriented blocks.
\end{definition}
\sergey{do we need to mention explicitly that blocks have the same orientation and are in the same table?}
We will sometimes write that $B'$ is a \textit{subblock} of $B$ or that $B$ is a \textit{superblock} of $B'$.
%
An example is that $T_1[:,\rangeto{3}{6}] \sqsubseteq T_1[:,\rangeto{3}{8}] $, which contains the sales numbers of all employees for the four quarters.

@@

\item the \textit{\CSignature} defines the properties that the arguments of the predicate must satisfy. This can be any property of individual blocks, such as type, length, size and orientation, as well as relations between properties of arguments, for example that the corresponding blocks must belong to the same table, or have equal type or length. In terms of logical and relational learning \cite{luc_book}, the \CSignature is known as the {\em bias} of the learner, it specifies when the constraint its arguments are well-formed.

@@

\begin{example}
Several constraint templates are illustrated in Table \ref{table:constraints}. A non-trivial example is the constraint template for the row-based sum:

@@

  \item \CSignature: $\sg_r$ has to be a single vector ($\textit{size}=1$) while $\mathbf{\sg_x}$ can be a block ($\textit{size}>=1$), which can be derived from the use of a normal or \textbf{bold} font. The two blocks have to be numeric. This constraint is orientation-specific, so it requires that the number of rows in $\mathbf{\sg_x}$ equals the length of $\sg_r$. While not strictly needed, we also add to the bias of this template that the number of columns to sum over is larger than $2$ as a block with two columns will also be captured by the $\eccalc{\sg_r}{\sg_1 + \sg_2}$ constraint;

@@

It is helpful to see the analogy of constraint templates with first order logic (FOL) and constraint satisfaction.
From a FOL perspective, the name of the constraint (e.g. \textit{SUM}) is just the predicate name, and the arguments $B_r$ and $B_x$ are the terms, which can be seen as either uninstantiated variables or as concrete values.
This also holds in our setting, where an instantiation of a variable corresponds to a concrete block. For example for $\ecrank{B_r}{B_x}$ and the spreadsheet in Figure~\ref{fig:main_example}, when we write $\ecrank{\range{T_1}{\rangeall}{8}}{\range{T_1}{\rangeall}{7}}$, then the value of $B_r$ is the $8$th vector in $T_1$: $B_r = \range{T_1}{\rangeall}{8} = [2,1,3,4]$ and the value of $B_x$ is the $7$th vector: $B_x = \range{T_1}{\rangeall}{7} = [1514, 1551, 691, 392]$.

@@

\newcommand{\sigc}{\ensuremath{\format{Sig}_c}}
\newcommand{\defc}{\ensuremath{\format{Def}_c}}
%\tias{TODO: conceptually $c(A_1, ..., A_n) :- Sig_c(A_1, ..., A_n) \wedge Def_c(A_1, ..., A_n)$ where '$c$' is here both the constraint name and the template}.
With this interpretation, we can speak about the signature and definition of a constraint template being \textit{satisfied}. We say that a signature (definition) of a constraint template with $n$ arguments is satisfied by the blocks $(B_1, ..., B_n)$ if $\sigc(B_1, ..., B_n)$ (respectively $\defc(B_1, ..., B_n)$) is satisfied. Likewise, the template is satisfied if both the signature and definition are satisfied; in logic programming, we would write a Prolog-like clause: $c(B_1, ..., B_n) \leftarrow \sigc(B_1, ..., B_n) \wedge \defc(B_1, ..., B_n)$. Under this interpretation, the term constraint and constraint template can be used interchangeably.

@@

      & $\numeric(\sg_{x})$ and $\ecperm{\sg_{x}}$

@@

We define the {\bf Tabular Constraint Learning Problem} as follows:
%
\begin{definition} \textit{Tabular Constraint Learning.}\label{def:problem_statement}\\
{\bf Given} a set of instantiated blocks ${\cal B}$ and a set of {\template}s ${\cal S}$: {\bf find} all constraints $c(B'_1, ..., B'_n)$ where $c$ is a constraint with a corresponding template in ${\cal S}$ and $(B'_1, ..., B'_n)$ is a valid argument assignment of the constraint $c$.
\end{definition}
% Let us now define what it means for a constraint from \template to hold in general. Let $t$ be a \template, $S$ be a mapping from variables to the subgroups of \groups, and $C$ be the constraint $t^S$, then $C$ \textit{holds} iff all constraints in $\text{\CSignature}^S$ and $\text{\CFunction}^S$ of $C$ hold.
% \sergey{moved from formalization END}

@@

\paragraph{Dependencies}
\sergey{question: the guidlines say caption for tables above the tables, but dsaa papers from 2014 and 2015 (I have checked three) put it below, that's confusing...}
In Table~\ref{table:constraints} one can see that for some constraints we used the predicate of another constraint in its signature, e.g. for \textit{PERMUTATION}. This expresses a dependency of the constraint on that other constraint. This can be interpreted as follows: the signature of the constraint consists of its own signature plus the signature of the depending constraint, and its definition of its own definition plus the definition of the depending constraint.
In FOL, we can see that one constraint entails the other, for example if $\ecperm{\sg_x}$ holds for a block $\sg_x$, then $\ecalldiff{\sg_{x}}$ also holds.

Apart from easing the specification of the signature and definition, in Section~\ref{sec:approach} we will see how such dependencies can be used to speed up the search for constraints.

@@

\subsubsection{Candidate group generation}
%The goal of this step is to prune out impossible groups from the beginning by looking at their properties rather than their actual content (\CSignature vs \CFunction).
Given a constraint template and the set of all groups ${\cal G}$, the goal is to find all combinations of groups that satisfy the constraint signature. Let there be $n$ groups and let the constraint have $k$ arguments, then a simple generate-and-test method would have to test $\tbinom{n+k-1}{k}$ combinations (given that groups can be repeated), which grows very rapidly with $n$.

However, the properties defined in the signature, such as related to type or size, can be used to avoid considering certain groups for certain arguments.
Furthermore, the choice of one argument can influence the possible candidates for the arguments, for example if they must have the same length.
Instead of writing specialized code for each of these cases, we make use of the built-in reasoning mechanisms of constraint satisfaction solvers.

A Constraint Satisfaction Problem (CSP) is a triple $(V,D,C)$ where $V$ is a set of \textit{variables}, $D$ the domain of possible values each variable can take and $C$ the set of constraints over $V$ that must be satisfied. In our case, we define one variable $V_i$ for each argument of the constraint. Each variable can be assigned to each of the groups in ${\cal G}$, so the domain consists of $|\{\cal G\}|$ identifiers.

@@

To reason over the groups, we automatically add to the constraint program a set of background facts that contain the properties of each group, namely its type, size, number of columns and rows, length, orientation, and table. We also add all previously found constraints of constraint templates that are implied by this template. The actual constraints then express the properties defined in the signature, such as whether arguments are of a certain type, have certain dependencies or have to be equal length or orientation, etc.
% \tias{More detailed description of the properties commented out, feel free to put in}
%Translation:
%\begin{description}
%  \item[Type] Require group that has supertype (e.g. integer => require numeric group, because subgroup might be integer)
%  \item[Length] Require group to have exact length (subgroups have the same length)
%  \item[Vectors / columns / rows] Require group to have at least the amount of vectors, columns, rows required
%  \item[Dependency] Fix dependent variables, for every assignment construct CSP that attempts to find values for the other variables
%  \item[Same length / table / orientation] Require groups to gave the same length table ...
%\end{description}

Constraints derived from the signature can usually but not always be enforced exactly on a group, as we have to consider whether subgroups could satisfy them.
Requirements on the lengths, orientations and tables of groups can be directly enforced since they stay the same for subgroups.
Typing requirements need to be relaxed to consider only non-subtypes, since, for example, numeric groups might contain integer subgroups.
Minimum sizes can be directly enforced, but exact size requirements are relaxed to minimum sizes, since subgroups might contains fewer vectors.
Finally, exact restrictions on the number of rows or columns may have to be relaxed depending on the orientation of the group.
Subgroups of row-oriented (column-oriented) group will always have the same number of columns (rows), however, the number of rows (columns) might decrease.
Therefore, we can only require row-oriented (column-oriented) groups to contain \textit{at least} the required amount of rows (columns).

For example, the \CSignature of ?? requires $v_r$ to be integer.
If a numeric group is assigned to $v_r$ then some subgroups may have type float while other may be of type integer, therefore, this group cannot be discarded.

\paragraph{Example}
We show this construction of a CSP for the constraint template $\ecsumprod{G'}{H'}{I'}$ in Table~\ref{table:constraints}. It defines that $G'$ is a cell whose values is the result of the sum of products of rows in $H'$ and columns in $I'$.
%
There are three \textit{variables} $X, Y, Z$ corresponding to group $G'$, $H'$ and $I'$ respectively. The domains are $D(X) = D(Y) = D(Z) = \{1, \ldots, |{\cal G}|\}$. Finally, the constraints are:
\begin{align*}
&numeric(X) \land numeric(Y) \land numeric(Z) \\
&\pcols(X) = 1 \land \prows(X) = 1 \\
&\plength(Y) = \plength(Z) \land \plength(Y) \geq 2
\end{align*}
\tias{Sumproduct is a rather peculiar constraint for spreadsheets... perhaps switch to one with a dependency, e.g. look-up?}
\sergey{ok, we need to see if look-up would fit here}
\\\\

The $\generategroups(\textit{t,{\cal G},C})$ method will construct this CSP and query a solver for all satisfying solutions. This is the required result set.

@@

Let us illustrate this on one of the spreadsheets from the gathered dataset. In Figure \ref{fig:autocompletion_example}, we see two tables the upper one is completed and in the lower one values in the last raw are not yet filled, then the system finds a satisfying constraint (\textit{SUMIF}) and based on it proposes filling values.

>>>>

\usepackage{times}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{xspace}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[numbers]{natbib}
\usepackage{notoccite}
\usepackage{framed}
\usepackage{amsmath}
\usepackage{tabularx}
\usepackage{subcaption}
\usepackage[inline]{enumitem}

@@

\newcommand{\CName}{Syntax\xspace}
\newcommand{\CSignature}{Signature\xspace}
\newcommand{\CFunction}{Definition\xspace}
\newcommand{\dependencies}{\ensuremath{\mathcal{D}}\xspace}
\newcommand{\groups}{\ensuremath{\mathcal{B}}\xspace}
\newcommand{\blocks}{\ensuremath{\mathcal{B}}\xspace}

@@

\newcommand{\numeric}{\format{numeric}}
\newcommand{\textual}{\format{textual}}
\newcommand{\integer}{\format{integer}}
\newcommand{\discrete}{\format{discrete}}
\newcommand{\plength}{\format{length}}
\newcommand{\psize}{\format{size}}
\newcommand{\ptype}{\format{type}}
\newcommand{\ptable}{\format{table}}
\newcommand{\por}{\format{orientation}}
\newcommand{\prows}{\format{rows}}
\newcommand{\pcols}{\format{columns}}
\newcommand{\nat}{\mathcal{N}}

\newcommand{\sg}{G'}

@@

The question that we answer in this paper is: is it possible to discover or reconstruct structural constraints (relations, functions) in flat tabular spreadsheet data?
To answer this question we contribute a general-purpose method and system, named \sname (pronounced as the word ``tackle''), for discovering row-wise and column-wise constraints.
%Furthermore, the algorithm accepts declarative specifications of the constraints and predicates of interest (the "bias"). \tias{revise declarative claim?}
It operates directly on (headerless) tables of a spreadsheet in an unsupervised setting, as it reasons on that raw tabular data directly with no example constraint instantiations given. %, and discovers column- and row-wise constraints.
We demonstrate the utility of our approach in an experimental evaluation.
Moreover, we sketch additional application scenario's such as autocompletion and error detection.
\\\\
This paper is organized as follows.
Section \ref{sec:formalization} introduces concepts relevant to our approach. Section \ref{sec:problem_statement} presents the problem statement. Section \ref{sec:approach} describes the approach. Section \ref{sec:evaluation} presents the evaluation of the approach. Section \ref{sec:applications} shows how our system can be used for applications. Section \ref{sec:related_work} presents an overview of the related work. Section \ref{sec:conclusions} provides conclusions.

@@

A row or columns is \textbf{type-consistent} if all cells in that row or column are of the same base type, that is, numeric or textual. We will use notation $T[a,{:}]$ to refer to the $a$-th row of table $T$, and similarly $T[{:},a]$ for the $a$-th column. For example in Figure~\ref{fig:main_example}, $T_1[1,:] = [1,2,3,4]$ and $T_5[:,1] = ['$Diana Coolen$', 5]$. The latter is not type-consistent while the former is.
%We will write \textbf{vector} to refer to a single row or column when its orientation does not matter.

@@

\begin{definition}
\textbf{Block containment $\sqsubseteq$.} 
A block $B'$ is contained in a block $B$, $B' \sqsubseteq B$, iff both are valid blocks (single orientation, contiguous, type consistent) and each of the vectors in $B'$ is also in $B$. For row-oriented blocks: $B' \sqsubseteq B \Leftrightarrow B=T[\rangeto{a}{b},:] \wedge B'=T[\rangeto{a'}{b'},:] \wedge a \leq a' \wedge b' \leq b$ and similarly for column-oriented blocks.
\end{definition}
We will sometimes write that $B'$ is a \textit{subblock} of $B$ or that $B$ is a \textit{superblock} of $B'$.
%
An example is that $T_1[:,\rangeto{3}{6}] \sqsubseteq T_1[:,\rangeto{3}{8}] $, which contains the sales numbers of all employees for the four quarters.

@@

\item the \textit{\CSignature} defines the properties that the arguments of the predicate must satisfy. This can be any property of individual blocks, such as type, length, size and orientation, as well as relations between properties of arguments, for example that the corresponding blocks must belong to the same table, or have equal type or length. In terms of logical and relational learning \cite{luc_book}, such a \CSignature is known as the {\em bias} of the learner, it specifies when the constraint its arguments are well-formed.

@@

\begin{example}
Several example constraint templates are illustrated in Table \ref{table:constraints}. An non-trivial example is the constraint template for the row-based sum:

@@

  \item \CSignature: $\sg_r$ has to be a single vector ($size=1$) while $\mathbf{\sg_x}$ can be a block ($size>=1$), which can be derived from the use of a normal or \textbf{bold} font. The two blocks have to be numeric. This constraint is orientation-specific, so it requires that the number of rows in $\mathbf{\sg_x}$ equals the length of $\sg_r$. While not strictly needed, we also add to the bias of this template that the number of columns to sum over is larger than $2$ as a block with two columns will also be captured by the $\eccalc{\sg_r}{\sg_1 + \sg_2}$ constraint;

@@

It is helpful to see the analogy of constraint templates with first order logic (FOL) and constraint satisfaction.
From a FOL perspective, the name of the constraint (e.g. \textit{SUM}) is just the predicate name, and the arguments $B$ and $A$ are the terms, which can be seen as either uninstantiated variables or as concrete values.
This also holds in our setting, where an instantiation of a variable corresponds to a concrete block. For example for $\ecrank{B_r}{B_x}$ and the spreadsheet in Figure~\ref{fig:main_example}, when we write $\ecrank{\range{T_1}{\rangeall}{8}}{\range{T_1}{\rangeall}{7}}$, then the value of $B_r$ is the $8$th vector in $T_1$: $B_r = \range{T_1}{\rangeall}{8} = [2,1,3,4]$ and the value of $B_x$ is the $7$th vector: $B_x = \range{T_1}{\rangeall}{7} = [1514, 1551, 691, 392]$.

@@

\newcommand{\sigc}{\ensuremath{\format{Sig}_c}}
\newcommand{\defc}{\ensuremath{\format{Def}_c}}
%\tias{TODO: conceptually $c(A_1, ..., A_n) :- Sig_c(A_1, ..., A_n) \wedge Def_c(A_1, ..., A_n)$ where '$c$' is here both the constraint name and the template}.
With this interpretation, we can speak about the signature and definition of a constraint template being \textit{satisfied}. We say that a signature (definition) of a constraint template with $n$ arguments is satisfied by the blocks $(B_1, ..., B_n)$ if $\sigc(B_1, ..., B_n)$ (respectively $\defc(B_1, ..., B_n)$) is satisfied. Likewise, the template is satisfied is both the signature and definition is satisfed; in first order logic, we would write: $c(B_1, ..., B_n) :-\sigc(B_1, ..., B_n) \wedge \defc(B_1, ..., B_n)$. Under this interpretation, the term constraint and constraint template can be used interchangeably.

@@

      & $\integer(\sg_{x})$ and $\ecperm{\sg_{x}}$

@@

We define the {\bf Tabular Constraint Learning Problem} as follows:
%
\begin{definition} \textit{Tabular Constraint Learning.}\label{def:problem_statement}\\
{\bf   Given} a set of instantiated blocks ${\cal B}$ and a set of {\template}s ${\cal S}$: {\bf Find} all constraints $c(B'_1, ..., B'_n)$ where $c$ is a constraint with a corresponding template in ${\cal S}$ and $(B'_1, ..., B'_n)$ is a valid argument assignment of the constraint $c$.
\end{definition}
% Let us now define what it means for a constraint from \template to hold in general. Let $t$ be a \template, $S$ be a mapping from variables to the subgroups of \groups, and $C$ be the constraint $t^S$, then $C$ \textit{holds} iff all constraints in $\text{\CSignature}^S$ and $\text{\CFunction}^S$ of $C$ hold.
% \sergey{moved from formalization END}

@@

\paragraph{Dependencies}
In Table~\ref{table:constraints} one can see that for some constraints we used the predicate of another constraint in its signature, e.g. for \textit{PERMUTATION}. This expresses a dependency of the constraint on that other constraint. This can be interpreted as follows: the signature of the constraint consists of its own signature plus the signature of the depending constraint, and its definition of its own definition plus the definition of the depending constraint.
In FOL, we can see that one constraint entails the other, for example $\ecperm{\sg_x} \models \ecalldiff{\sg_{x}}$.

Apart from easing the specification of the signature and definition, in Section~\tias{TODO} we will see how such dependencies can be used to speed up the search for constraints.

@@

\subsubsection{Generating superblock assignments}
%The goal of this step is to prune out impossible groups from the beginning by looking at their properties rather than their actual content (\CSignature vs \CFunction).
Given a constraint template~$s$ and the set of all blocks~\blocks, the goal is to find all combinations of blocks that satisfy the constraint signature. Let there be $n$ blocks and let~$s$ have~$k$ arguments, then a simple generate-and-test method would have to test $n^k$ combinations (given that blocks can be repeated), which can grow very rapidly with $n$, depending on $k$.

However, the properties defined in the signature, such as those related to the type or size of a block, can be used to avoid considering certain groups for certain arguments.
Furthermore, the choice of one argument can influence the possible candidates for the other arguments, for example, if they must have the same length.
Instead of writing specialized code for each of these cases, we make use of the built-in reasoning mechanisms of constraint satisfaction solvers.

A Constraint Satisfaction Problem (CSP) is a triple $(V,D,C)$ where $V$ is a set of \textit{variables}, $D$ the domain of possible values each variable can take and $C$ the set of constraints over $V$ that must be satisfied.
In our case, we define one variable $V_i$ for each argument of a constraint template.
Each variable can be assigned to each of the blocks in \blocks, so the domain consists of $|\{\blocks\}|$ identifiers.

@@

To reason over the blocks, we automatically add to the constraint program a set of background facts that contain the properties of each block, namely its type, size, number of columns and rows, length, orientation, and table.
\samuel{Review: We also add all previously found constraints of constraint templates that are implied by this template.}
The actual constraints correspond to the requirements defined in the signature, such as whether arguments are of a certain type, have certain dependencies or have to be equal length or orientation, etc.

% \tias{More detailed description of the properties commented out, feel free to put in}
%Translation:
%\begin{description}
%  \item[Type] Require group that has supertype (e.g. integer => require numeric group, because subgroup might be integer)
%  \item[Length] Require group to have exact length (subgroups have the same length)
%  \item[Vectors / columns / rows] Require group to have at least the amount of vectors, columns, rows required
%  \item[Dependency] Fix dependent variables, for every assignment construct CSP that attempts to find values for the other variables
%  \item[Same length / table / orientation] Require groups to gave the same length table ...
%\end{description}

\begin{table}[htb]
  \caption{Translation of signature requirements to superblock constraints}
  \label{tbl:translation}
  \begin{tabularx}{\linewidth}{lX}
    \textbf{Constraint} & \textbf{Superblock constraint} \\ \hline \hline
    $\plength(B) = x$ & $\plength(B) = x$ \\ \hline
    $\por(B) = x$ & $\por(B) = x$ \\\hline
    $\ptable(B) = x$ & $\ptable(B) = x$ \\ \hline
    $\ptype(B) = x$ & $\mathit{supertype}(\ptype(B)) = \mathit{supertype}(x)$ \\ \hline
    $\psize(B) \geq x$ & $\psize(B) \geq x$ \\ \hline
    $\psize(B) = x$ & $\psize(B) \geq x$ \\ \hline
    $\pcols(B) = x$ & $\mathit{if}~\por(B) = \mathit{column}: \pcols(B) \geq x$ \\ 
    & $\mathit{if}~\por(B) = \mathit{row}: \pcols(B) = x$ \\ \hline
    $\prows(B) = x$ & $\mathit{if}~\por(B) = \mathit{column}: \prows(B) = x$ \\ 
    & $\mathit{if}~\por(B) = \mathit{row}: \prows(B) \geq x$
  \end{tabularx}
\end{table}

In order for our approach to be complete, the signature constraints can not always be enforced directly on a block.
If a block matches the signature it is a valid candidate, however, if it does not there might exist subblocks that still satisfy the signature.
\begin{example}
  Consider, for example, the constraint template \ecseries{x} and a superblock candidate~$B$ for~$x$ containing both integer and floating point values.
  While the signature of the template requires~$x$ to be integer we cannot discard~$B$ since there might be subblocks~$B' \sqsubseteq B$ that only contain integer values and satisfy the signature.
\end{example}
Table~\ref{tbl:translation} shows the how constraints from the signature are converted to constraints on superblocks.
Requirements on the lengths, orientations and tables of blocks can be directly enforced since they are invariant under block containment($\sqsubseteq$).
Typing requirements need to be relaxed to check only for supertypes (numeric and textual).
Minimum sizes can be directly enforced, but exact size requirements are relaxed to minimum sizes, since subgroups might contains fewer vectors.
Finally, restrictions on the number of rows or columns behave as length or size constraints based on the orientation of the group they are applied to.
Subgroups of row-oriented (column-oriented) group will always have the same number of columns (rows), however, the number of rows (columns) might decrease.

The $\generategroups(\textit{s,\blocks})$ method will use these conversion rules to construct a CSP and query a solver for all satisfying solutions.
These solutions will correspond with the valid superblock assignments for constraint template~$s$.

\begin{example}
  \samuel{Is this still necessary?}
  Consider the template $\ecsumprod{x}{y}{z}$, then the generated CSP will contain three variables $V_x, V_y, V_z$ corresponding to arguments $x, y$ and $z$ respectively.
  Given blocks~\blocks, the domain of these variables are $D(V_x) = D(V_y) = D(V_z) = \{ 1, \dots, |\blocks| \}$.
  Finally, the signature will be translated into constraints:
  \begin{align*}
    & \numeric(V_x) \land \numeric(V_y) \land \numeric(V_z) \\
    & \por(V_x) = \mathit{row} \implies \pcols(V_x) = 1 \\
    & \por(V_x) = \mathit{row} \implies \prows(V_x) = 1 & \\
    & \plength(V_y) \geq 2 \land \plength(V_z) \geq 2 \\
    & \plength(V_y) = \plength(V_z)
  \end{align*}
  Every solution is a tuple of values for~$V_x$, $V_y$ and~$V_z$ corresponding directly with a superblock assignment over~\blocks.
\end{example}

@@

Let us illustrate this on one of the spreadsheets from the gathered dataset. In Figure \ref{fig:autocompletion_example}, \sergey{to be finished}
